Polynomial Time and Dependent Types

### 2023-10-24 paper/paper.tex
[ paper ] comment on proving 3-SAT is in NP

The realisability type $\mathbf{R}(A)$ described in
\autoref{sec:qtt-reflection} allows us to internalise the
realisability of certain functions into the logical ($\sigma = 0$)
fragment of the calculus. However, it is not possible to derive any
logical consequences from this other than turning it back into a
function. This limitation becomes acute when trying to prove results
from standard Computational Complexity theory. Even though we can
characterise the class NP, as we did in \autoref{sec:np-class}, we
cannot prove the Cook-Levin theorem that 3-SAT is NP-complete. This is
because the proofs relies on obtaining the \emph{source code} of the
program solving an NP problem and then encoding that program in
3-SAT. To do this in our setting, we would need to internalise the
soundness property (\thmref{thm:lfpl-soundness}) as an axiom, stating
that for a realisable polytime function there (merely) exists a
realising expression $E$ that completes in polynomial time, and then
writing polytime enodings into 3-SAT. We hope that the addition of
such an axiom to our system would lead to an expressive machine-free
\emph{Synthetic Computational Complexity Theory}, analogous to the
Church-Turing axiom for Synthetic Computability Theory \cite{Bauer06}.

>>>>

The realisability type $\mathbf{R}(A)$ described in
\autoref{sec:qtt-reflection} allows us to internalise the
realisability of certain functions into the logical ($\sigma = 0$)
fragment of the calculus. However, it is not possible to derive any
logical consequences from this other than turning it back into a
function. This limitation becomes acute when trying to prove results
from standard Computational Complexity theory. Even though we can
characterise the class NP, as we did in \autoref{sec:np-class}, and it
is a ``matter of programming'' to show that 3-SAT is in NP, we cannot
prove the Cook-Levin theorem that 3-SAT is NP-complete. This is
because the proofs relies on obtaining the \emph{source code} of the
program solving an NP problem and then encoding that program in
3-SAT. To do this in our setting, we would need to internalise the
soundness property (\thmref{thm:lfpl-soundness}) as an axiom, stating
that for a realisable polytime function there (merely) exists a
realising expression $E$ that completes in polynomial time, and then
writing polytime enodings into 3-SAT. We hope that the addition of
such an axiom to our system would lead to an expressive machine-free
\emph{Synthetic Computational Complexity Theory}, analogous to the
Church-Turing axiom for Synthetic Computability Theory \cite{Bauer06}.
